---
title: "Reproducible Research"
author: "Kassim"
date: "8/16/2020"
output:
  html_document: default
  word_document: default
---

```{r}
act_data <- read.csv("activity.csv", header=TRUE, sep=",")
str(act_data)
```

## R Markdown

Process the data

```{r}
act_data$date <- as.Date(act_data$date, format="%Y-%m-%d")
act_data$interval <- as.factor(act_data$interval)
str(act_data)
head(act_data, 10)
steps_per_day <- aggregate(steps ~ date, data=act_data, FUN=sum)
colnames(steps_per_day) <- c("date", "steps")
```

## Including Plots

Plotting the report

```{r echo=TRUE}
ggplot(steps_per_day, aes(x = steps)) + 
  geom_histogram(fill = "blue", binwidth = 1000) + 
  labs(title = "Histogram - Steps Taken Per Day", x = "Steps Per Day", y = "Frequency")
```
Report the mean and median of the total number of steps taken per day
```{r echo=TRUE}
mean_steps_per_day <- mean(steps_per_day$steps)
mean_steps_per_day
median_steps_per_day <- median(steps_per_day$steps)
median_steps_per_day
```
Average daily activity pattern
```{r echo=TRUE}
steps_per_interval <- aggregate(steps ~ interval, data = act_data, FUN = mean, na.rm = TRUE)
steps_per_interval$interval <- as.integer(levels(steps_per_interval$interval)[steps_per_interval$interval])
colnames(steps_per_interval) <- c("interval", "steps")
```
```{r echo=TRUE}
ggplot(steps_per_interval, aes(x = interval, y = steps)) + 
  geom_line(col = "blue", size = 1) + 
  labs(title = "Average Daily Activity Pattern", x = "Interval", y = "Steps")

```
5-minute interval, on average across all the days in the dataset
```{r echo=TRUE}
max_interval <- steps_per_interval[which.max(steps_per_interval$steps),]
max_interval
```
Missing values
```{r echo=TRUE}
missing_values <- sum(is.na(act_data$steps))
missing_values
```
New dataset to fill in missing values
```{r echo=TRUE}
new_act_data <- act_data
index_of_na <- which(is.na(new_act_data$steps))
for (i in index_of_na) {
  new_act_data$steps[i] <- with(steps_per_interval, steps[interval = new_act_data$interval[i]])
}
```
Histogram of the total number of steps taken each day after missing values are imputed:
```{r echo=TRUE}
new_steps_per_day <- aggregate(steps ~ date, data = new_act_data, FUN=sum)
colnames(new_steps_per_day) <- c("date", "steps")
ggplot(new_steps_per_day, aes(x = steps)) + 
  geom_histogram(fill = "blue", binwidth = 1000) + 
  labs(title = "Histogram - Steps Taken Per Day", x = "Steps Per Day", y = "Frequency")
```
Activity patterns between weekdays and weekends:
```{r echo=TRUE}
dt <- data.table(new_act_data)
dt[, weekday := ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")]
dt$weekday <- as.factor(dt$weekday)
dt$interval <- as.integer(levels(dt$interval)[dt$interval])
head(dt, 10)
```
Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:
```{r echo = TRUE}
steps_per_weekday <- aggregate(steps ~ interval+weekday, data = dt, FUN = mean)
ggplot(steps_per_weekday, aes(x = interval, y = steps)) + 
  geom_line(col = "blue", size = 1) + 
  facet_wrap(~ weekday, nrow=2, ncol=1) + 
  labs(x = "Interval", y = "Number of Steps")
```
